### 2014-03-17 reanalysis/Fadi_Towfic_generic_medicine/towfic_et_al_tomuch.rmd

hist(fstatistics[table_s4$Probe], breaks=100, add=F, freq=T, border=thiscolors[2])
```
The scatter plots of the values shows a weak correlation, but they are all high. And the histogram shows that the probes from Table S4 all have a high reproduced F-statistic compared to the overall.

@@

How this fold change was calculated is not explained. A common way is to take the difference of geometric means that were calculated for Table S2 above.
```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
foldchange_DP_N = meanDP - meanN

@@

# plot
plot(table_s5$Fold_Change, foldchange_DP_N[table_s5$Probe],
     ylim=lim, xlim=lim, pch=".", xlab="Table_S5", ylab="reproduced",
     main="FC comparison")

```
The plot reveals that the reproduced fold changes are much more compressed than the ones in Table S5.
There seems to two linear relationship, one for neagtive FC's and another for positve FC's. They seems to have different intercepts but comparable slope.
The methods part is more specific;
>  First, we computed the statistical significance of differential expression between treatments using the Analysis of Variance (ANOVA) method for each probe [21], adjusting for multiple hypothesis testing using the Benjamini-Hochberg False Discovery Rate (FDR) correction [22].

#"Probes with Fold Change <  1.10 notdisplayed"

# Table s5 seems to be a comparison of expression in  GA (drug product) vs. generic (N) , listing the probes with greatest fold change (>abs(1.1)??) and the p-value calculated by several different tests. There seems to be no description of how the fold-change was calculated and after an inspection they look weird.

meanDP = rowMeans(thisdata[, sampleannotation$covariate=="DP"]) # DP is drug product also called GA
meanN = rowMeans(thisdata[, sampleannotation$covariate =="N"]) # N is Natco the generic,  (or G for alternative covariates)

# fold change between N and DP, calculated as geometric mean.
x = table_s5$Fold_Change
y = (meanDP - meanN)[table_s5$Probe]
lim =c( min(   c(min(x), min(y))  ), max( c(max(x), max(y)) ))
plot(x, y , ylim=lim, xlim=lim, pch=".", xlab="Table_S5", ylab="reproduced", main="FC comparison") # Not a perfect match. And the table_s5 numbers seems offset.
```
```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
# table S2 has the AVG for both DP and N listed
fc_table_s2 = table_s2$AVG_GA - table_s2$AVG_generic
names(fc_table_s2) = table_s2$ID
x = table_s5$Fold_Change
y = fc_table_s2[table_s5$Probe]
lim =c( min(   c(min(x, na.rm=T), min(y, na.rm=T))  ), max( c(max(x, na.rm=T), max(y, na.rm=T)) ))
plot(x, y , ylim=lim, xlim=lim, pch="." ,  xlab="Table_S5", ylab="Table_S2", main="FC comparison") # perfect match.
```
But this is odd, we were able to reproduce table s2 and table s2 matches table s5, but table s5 fold changes does not match our reproduced values as good as the match for table s2.

The reason that the line is not as straight for the reproduce FC's for table s5 as for the mean expressions in table S2 is because of different scale in the plot and the fact that the small difference will be higher for a ratio (combining to values with a individual deviation from the tables.)
```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
x = (meanDP - meanN)[table_s2$ID]
y = fc_table_s2
plot(x, y , ylim=lim, xlim=lim, pch="." ,  xlab="reproduced", ylab="Table_S2", main="FC comparison") # perfect match.
```

```{r}
# Still there is a offset in the numbers.
commonprobes = table_s5$Probe[table_s5$Probe %in% table_s2$ID]
fc_table_s5 = table_s5$Fold_Change[match(commonprobes ,table_s5$Probe)]
head(cbind(fc_table_s2[commonprobes], fc_table_s5, fc_table_s5-fc_table_s2[commonprobes], fc_table_s5/fc_table_s2[commonprobes]))
# seems that some unacounted transformation has occured on the fold changes.
```
```{r}
source("../../commonscripts/helper_functions.r")
getdifftab_limma
```
```{r}
# reproduce the Nom P-value columns
limma_ret= getdifftab_limma(edata=datamatrices[["real_combat_covariates"]], condition=sampleannotation[,"covariate"], contrast="DP-N" )
limma_p = limma_ret$p
names(limma_p) = limma_ret$gene
DP = sampleannotation$covariate=="DP"
N = sampleannotation$covariate=="N"
ttest_p = apply(datamatrices[["real_combat_covariates"]], 1, function(x) { t.test(x[DP], x[N])$p.value } )
anova_p = apply(datamatrices[["real_combat_covariates"]], 1, function(x){ summary(aov(x[DP|N]~sampleannotation$covariate[DP|N]))[[1]][1,5]}  )  # aov er kanskje feil for dette, Er ikke live mange pr?ver i guppene.

# wilcox.test gives a few hundred warnings for ties.
# Warning: cannot compute exact p-value with ties
wilcoxon_p = suppressWarnings(apply(datamatrices[["real_combat_covariates"]], 1, function(x) { wilcox.test(x[sampleannotation$covariate=="DP"], x[sampleannotation$covariate=="N"])$p.value } )
)
```
```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
par(mfrow=c(2, 2))
#plot(table_s5$Fold_Change, (meanDP - meanN)[table_s5$Probe] ,  ylim=c(-1.5,1.5), xlim=c(-1.5,1.5), pch=".", xlab="Table_S5", ylab="reproduced", main="FC comparison")
plot(log10(table_s5$t_test_p), log10(ttest_p[table_s5$Probe]), pch=".", main="t-test p")
plot(log10(table_s5$limma_p), log10(limma_p[table_s5$Probe]), pch=".", main="limma p")
plot(log10(table_s5$anova_p), log10(anova_p[table_s5$Probe]), pch=".", main="anova p")
plot(log10(table_s5$wilcoxon_p), log10(wilcoxon_p[table_s5$Probe]), pch=".", main="wilcoxon p")
```
# The p-values does not make a straight line, but they are low. But maybe all are low?

>>>>

#hist(fstatistics[table_s4$Probe], breaks=100, add=F, freq=T, border=thiscolors[2])
```
The scatter plots of the values shows a weak correlation, but they are all high. And the histogram shows that the probes from Table S4 all have a high reproduced F-statistic compared to the overall.

@@

There must be a typo in the figure text, they ment "Comparison of expression in GA to expression in **generic** for each probe, ...".

> To find differentially expressed probes between generic and GA, we utilized various statistical tests at the probe level and merged the results across the different methods. <cite> Towfic et al.

How the fold change was calculated is not explained. A common way is to take the difference of geometric means that were calculated for Table S2 above.
```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
foldchange_DP_N = meanDP - meanN

@@

# plot
plot(table_s5$Fold_Change, foldchange_DP_N[table_s5$Probe],
     ylim=lim, xlim=lim, pch=".", xlab="Table_S5", ylab="reproduced",
     main="FC comparison")
rm(lim)
```
The plot reveals that the reproduced fold changes are much more compressed than the ones in Table S5.
There seems to two linear relationship, one for neagtive FC's and another for positve FC's. They seems to have different intercepts but comparable slope.

The average expression for "GA" (DP) and "generic" (N) are also given in Table S2. The fold change calculated from those numbers compared to the ones given in Table S5 shows a similar pattern:

```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
# table S2 has the AVG for both DP and N listed
fc_table_s2 = table_s2$AVG_GA - table_s2$AVG_generic
names(fc_table_s2) = table_s2$ID
x = table_s5$Fold_Change
y = fc_table_s2[table_s5$Probe]
lim =c( min(   c(min(x, na.rm=T), min(y, na.rm=T))  ), max( c(max(x, na.rm=T), max(y, na.rm=T)) ))
plot(x, y , ylim=lim, xlim=lim, pch="." ,  xlab="Table_S5", ylab="Table_S2",
     main="FC comparison, given values.") #  almost perfect match with offset.
rm(x, y, lim, fc_table_s2)
```
We are not able to reporduce the foldchanges fully.

>  First, we computed the statistical significance of differential expression between treatments using the Analysis of Variance (ANOVA) method for each probe [21]...<cite> Towfic et al.

```{r}
anova_p = apply(datamatrices[["real_combat_covariates"]], 1,
              function(x){ summary(aov(x[DP|N]~sampleannotation$covariate[DP|N]))[[1]][1,5]})
```

> Next, we utilized Linear Models for Microarray (LIMMA) data analysis [23], [24] R package, part of the Bioconductor framework [25], to compare generic and GA samples, fitting a linear model that adjusts for fixed effect from medium (Effect = (GA-generic) – (generic-Medium)). The coefficients for the linear model were tested for significance using a modified t-test (taking into account standard deviation)...<cite> Towfic et al.

The formula "Effect = (GA-generic) – (generic-Medium)" seems strange. A more standard way would be GA-generic.

```{r}
group = factor(sampleannotation$covariate)
design = model.matrix(~0 + group)
fit = lmFit(datamatrices[["real_combat_covariates"]], design)
cont.matrix = makeContrasts ( contrasts="(groupDP-groupN) - (groupN-groupM)", levels=design)  
fit2 = contrasts.fit(fit, cont.matrix)
limma_p_org = eBayes(fit2)$p.value

cont.matrix = makeContrasts ( contrasts="groupDP-groupN", levels=design)  
fit2 = contrasts.fit(fit, cont.matrix)
limma_p_alt = eBayes(fit2)$p.value
```
> In parallel, we used Comparative Marker Selection as implemented in GenePattern [26] to directly compare probes between generic and GA. We applied two separate techniques within this framework; a traditional T-test and a Signal-to-Noise Ratio test (SNR).

The T-test is stright forward.
```{r}
DP = sampleannotation$covariate=="DP"
N = sampleannotation$covariate=="N"
ttest_p = apply(datamatrices[["real_combat_covariates"]], 1, function(x) { t.test(x[DP], x[N])$p.value } )
```

> ...we used the Wilcoxon Rank Sum Test [27] as implemented in R ...

```{r}
# wilcox.test gives a few hundred warnings for ties:
# Warning: cannot compute exact p-value with ties
wilcoxon_p = suppressWarnings(apply(datamatrices[["real_combat_covariates"]], 1, 
                                    function(x) { wilcox.test(x[sampleannotation$covariate=="DP"],
                                                              x[sampleannotation$covariate=="N"])$p.value } )
)
```

```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
par(mfrow=c(2, 2))
#plot(table_s5$Fold_Change, (meanDP - meanN)[table_s5$Probe] ,  ylim=c(-1.5,1.5), xlim=c(-1.5,1.5), pch=".", xlab="Table_S5", ylab="reproduced", main="FC comparison")
plot(log10(table_s5$t_test_p), log10(ttest_p[table_s5$Probe]), pch=".", main="t-test p")
plot(log10(table_s5$limma_p), log10(limma_p[table_s5$Probe]), pch=".", main="limma p")
plot(log10(table_s5$anova_p), log10(anova_p[table_s5$Probe]), pch=".", main="anova p")
plot(log10(table_s5$wilcoxon_p), log10(wilcoxon_p[table_s5$Probe]), pch=".", main="wilcoxon p")
```
# The p-values does not make a straight line, but they are low. But maybe all are low?
